---
title: "EuroCrime Stats 2015 Analysis with R"
author: "Shivam Saxena"
date: "9 October 2018"
output:
  html_document: default
---

### Data Load
```{r }
library(knitr)
library(kableExtra)
euro_crimes_data = read.csv("EurostatCrime2015.csv",header = TRUE)
colnames(euro_crimes_data) <- c("Country","Assault",	"Intentional homicide",	"Rape",	"Robbery",	"Sexual assault",	"Sexual violence",	"Theft")
```

### Knowing the Size and Structure of Data
```{r}
dim(euro_crimes_data)
str(euro_crimes_data)
```

### Grouping Similar Columns Rape , Sexual Assault and Sexual Violence clubbed to New Column Sex.crime
```{r}
euro_crimes_data$Sex.crime <- euro_crimes_data$Rape + euro_crimes_data$`Sexual assault` + euro_crimes_data$`Sexual violence` 
kable(euro_crimes_data) %>%
  kable_styling(bootstrap_options = "striped", font_size = 10)
```

#### Removed Columns Rape , Sexual Assault and Sexual Violence 
```{r}
crimes_dat2 = subset(euro_crimes_data, select = -c(Rape,`Sexual assault`,`Sexual violence`) )
```


### Data Cleaning , checking rows having 'NA' values
```{r}
missing_Data = crimes_dat2[rowSums(is.na(crimes_dat2)) > 0,]
kable(missing_Data) %>%
  kable_styling(bootstrap_options = "striped", font_size = 12)
```

### 'NA' tuples removed
```{r}
crimes_dat_na_cleaned = crimes_dat2[complete.cases(crimes_dat2),]
kable(crimes_dat_na_cleaned) %>%
  kable_styling(bootstrap_options = "striped", font_size = 12)
```

### Cleaned DataFrame Size
```{r}
dim(crimes_dat_na_cleaned)
```

# Analysis

### Theft was the most common crime
```{r}
irl_crime_dat = subset(euro_crimes_data,euro_crimes_data["Country"] == "Ireland")
colnames(irl_crime_dat)[apply(irl_crime_dat,1,which.max)]

```

### least 3 common crimes are:Sexual assault , Rape	, Intentional homicide
```{r}
irl_crime_dat <- irl_crime_dat[1,2:9]
irl_crime_dat = sort(irl_crime_dat,decreasing = TRUE)

kable(irl_crime_dat[,6:8]) %>%
  kable_styling(bootstrap_options = "striped", font_size = 12)
```

### Highest record of offences - Sweden
```{r}
euro_crimes_data$total_offences <- rowSums(euro_crimes_data[2:9])
ordered_euro_crimes_data = euro_crimes_data[order(euro_crimes_data$total_offences,decreasing = TRUE,na.last = NA),]
kable(ordered_euro_crimes_data[1,]) %>%
  kable_styling(bootstrap_options = "striped", font_size = 12)

```

# Most Common Crime Across Europe

__A new row can be generated from the given data which will contain the total occurences of each crime type across all the countries.__ 
```{r}
total_euro_crimes_data <- colSums(euro_crimes_data[2:9],na.rm = TRUE)
total_euro_crimes_data <- data.frame(total_euro_crimes_data)
total_euro_crimes_data <- cbind( crime_types = rownames(total_euro_crimes_data), total_euro_crimes_data)
rownames(total_euro_crimes_data) <- 1:nrow(total_euro_crimes_data)
```

__It will help to interpret which crime is most common across Europe Continent.__  

__Generated Barplot:__  

```{r}
x <- total_euro_crimes_data$crime_types
y <- total_euro_crimes_data$total_euro_crimes_data

library(ggplot2)
# bar plot showing the crimes which are most occurred in europe
ggplot(data=total_euro_crimes_data, aes(x=crime_types, y=total_euro_crimes_data)) +
  geom_bar(stat="identity", fill="steelblue")+
   coord_flip() +
  geom_text(aes(label=total_euro_crimes_data), position=position_dodge(width=0.9), vjust=-0.25)
```

__From the plot it is clearly shown that "Theft" is the most common across EU while "Intentional Homicide" is least common.__
